---
title: "Code Main"
output: html_document
date: "2024-02-05"
---
R setup - packages
```{r setup, include=FALSE}
pkgs <- c("gsw", "tidyverse", "ncdf4", "lubridate", "dplyr", "this.path") # required packages
for(i in 1:length(pkgs)){
  # Load packages, installing if necessary
  j <- library(pkgs[i], character.only = TRUE, logical.return = TRUE)
  if(!j){
    install.packages(pkgs[i])
    library(pkgs[i], character.only = TRUE)}}
```

Import Data
```{r}
setwd(this.dir()) # set working directory
data.list <- list() # 'list' is a reserved keyword R uses to create lists -- avoid using reserved keywords to define objects
# list <- list()
data_files <- list.files(recursive = TRUE)
data_files <- data_files[grepl('\\.nc', data_files)]
for(i in 1:length(data_files)){
  nc_ds <- nc_open(data_files[i])
  dim_lat <- ncvar_get(nc_ds, "LATITUDE")
  dim_lon <- ncvar_get(nc_ds, "LONGITUDE")
  dim_temp <- ncvar_get(nc_ds, "TEMP")
  dim_psal <- ncvar_get(nc_ds, "PSAL")
  dim_pres <- ncvar_get(nc_ds, "PRES")
  dim_time <- ncvar_get(nc_ds, "TIME")
  
  dim_depth <- gsw_z_from_p(dim_pres, dim_lat) # calculate depth from pressure and latitude
  
  # origin <- as.Date('1950-01-01')
  origin <- '1950-01-01'
  sec <- dim_time * 24 * 60 * 60
  # sec <- dim_time[1] * 24 * 60 * 60
  dim_time <- as.POSIXct(sec, origin = origin) # convert stored 'time' into actual sample time
  dim_year <- as.numeric(strftime(dim_time, '%Y'))
  dim_month <- as.numeric(strftime(dim_time, '%m'))
  dim_day <- as.numeric(strftime(dim_time, '%d'))
  
  j <- which.min(apply(dim_depth, 2, function(z) min(z, na.rm = TRUE)))
  
  # name the variables for clarity, rather than using X1, X2 etc
  tesa <- data.frame(depth = dim_depth[,j], psal = dim_psal[,j], temp = dim_temp[,j], pres = dim_pres[,j])
  # tesa <- data.frame(cbind(dim_psal[,j], dim_temp[,j], dim_pres[,j]))
  tesa
  
  # Can combine new variables within a single call of 'mutate'
  tesa <- tesa %>% mutate(lon = dim_lon[j],
                          lat = dim_lat[j],
                          year = dim_year[j],
                          month = dim_month[j],
                          day = dim_day[j]
  )
  data.list[[i]] <- tesa
}

# Include a sample index in each list element
data.list <- lapply(1:length(data.list), function(z){
  x <- data.list[[z]]
  x$sample <- z
  x})


data.list
new.data <- do.call('rbind', data.list)
summary(na.omit(new.data$lat))


new.data$asal <- gsw_SA_from_SP(new.data$psal, p = new.data$pres,
                                longitude = new.data$lon, latitude = new.data$lat)
# Calculate seawater density
new.data$rho <- gsw_rho(SA = new.data$asal, CT = new.data$temp, p = new.data$pres)
new.data$rho <- new.data$rho / 1000 # convert units to g/cm^3
# we can convert back to a list, but it's probably easier to work with the data
# frame, using the sample index column to differentiate samples.
data.list <- lapply(unique(new.data$sample), function(z){
  new.data[new.data$sample == z,]})

#for(i in unique(new.data$sample)){
  #d <- new.data[new.data$sample == i,]
  #Title <- paste(d$day[1], month.abb[d$month[1]], d$year[1])
  #plot(d$rho, d$depth, xlab = expression(sewater ~ density ~ (g/cm^3)),
  #     ylab = 'depth (m)', main = Title)
#}
```
Not entirely sure what is wrong with the above code. It works just fine in a normal
R script, but returns an error when run through R Markdown. There are no other
issues running the rest of the Markdown script.

Interpolate density by depth
```{r}
# Average the density-at-depth measurements across samples within months. The depths
# may not line up, so first create smooth interpolating functions to find density
# at the same depths in every sample. Then create a smooth interpolating function
# that returns density-at-depth for each month (this is needed for our equations).
new.data$depth <- new.data$depth * 100 # convert units of depth to cm (our interpolating functions are used within the sinking speed equations which use g-cm-s units)
months <- unique(new.data$month)
nmonths <- length(months)
depth.range <- data.frame(min = rep(NA, nmonths), max = rep(NA, nmonths), month = months) # store range of sampled depths for each month
rho.interp <- setNames(vector(mode = 'list', length = nmonths), month.abb[months]) # store interpolating function for each month
for(i in 1:nmonths){
  d <- new.data[new.data$month == months[i],] # subset data for month i
  samples <- unique(d$sample) # sample indices for month i
  nsamples <- length(samples)
  dr <- sapply(samples, function(z){ # depth range of each sample
    x <- d[d$sample == z,]
    range(x$depth, na.rm = TRUE)})
  dr <- c(ceiling(min(dr[2,])), floor(max(dr[1,]))) # a depth range for month i  
  depth.range[depth.range$month == months[i], c('min','max')] <- dr
  depth.inc <- seq(dr[1], dr[2], -10) # uniform depth increments [cm] to feed into interpolating functions
  ndepth <- length(depth.inc)
  rho.matrix <- matrix(NA, ndepth, nsamples) # store interpolated densities for each sample
  for(j in 1:nsamples){
    dj <- d[d$sample == samples[j],] # subset data for sample j in month i
    f <- approxfun(dj$depth, dj$rho)
    rho.matrix[,j] <- f(depth.inc) # interpolate
  }
  rho <- rowMeans(rho.matrix, na.rm = TRUE) # average density (at depths depth.inc) across samples
  rho.interp[[i]] <- approxfun(depth.inc, rho, rule = 2) # create and store a single interpolating function for density-at-depth per month
}
```

Define average fecal pellet parameters and mathematical constants
```{r}
# Set up a numerical integration of the sinking & degradation equations.
# Choose a time step. The value matters -- there is a trade off between accuracy
# and time taken to solve the equations. A small time step will provide more
# accurate answers, but take longer (this can be 'dialed-in' later).
dt <- 10 * 60 # 10 minute time step [s]
# Set parameter values (in correct units)
L.init <- 0.2927 # initial length [cm]
D.init <- 0.0183 # initial diameter [cm]
gd <- 1.121 # faecal pellet density [g/cm^3]
# rD <- 1.2 / 86400 # degradation rate [1/s]
vis <- 0.0189 # seawater viscosity []
coef <- 0.0790 # scaling coefficient from Komar
g <- 981 # acceleration due to gravity [cm/s^2]
p <- -1.644 # power term [dimensionless] from Komar
# We know faecal pellet decay rates at two specific depths from Morata
rD_13 <- 0.07 / 86400 # degradation rate [1/s] at chlorophyll max (13 m)
rD_90 <- 0.02 / 86400 # degradation rate [1/s] at 90 m
# Convert decay rates into the Martin curve parameters
rD_b <- log(rD_90 / rD_13) / log(1300 / 9000) # power term [dimensionless]
rD_a <- rD_13 * 1300 ^ rD_b # coefficient [cm/s]: equivalently expressed as rD_a <- rD_90 * 9000 ^ pow

# Fragmentation rate of faecal pellets. Don't yet know a realistic value.
# This simple representation of fragmentation rate should be extended to account for
# faecal pellet size and extent of plastic contamination.
rF <- 0.05 / 86400 # fragmentation rate [1/s] ()

```

Functions
```{r}
source('volume function.R') # loads volume_fun
# volume <- function(L,D){
#   pi / 4 * L * D^2}
mass <- function(V, gd){
  V * gd}
shape <- function(L, D){
  L / D}

# mass.degrade <- function(t, r, m0 = m.init){
#   # This function models faecal pellets degrading at a constant rate determined by rD.
#   # This is an unrealistic assumption because rD changes with depth.
#   # Given the initial mass and rD, this function outputs mass at any time t.
#   m0 * exp(-r * t)}

rD_fun <- function(z, a = rD_a, b = rD_b){
  # Degradation rate [1/s] as a function of depth.
  # Martin curve (is just an exponential decay with depth)
  a * z ^ -b
}

speed <- function(m, rhoF, rhoW, visc = vis, 
                  grav = g, pow = p, const = coef, s = S){
  const / visc * {rhoF - rhoW} * grav * s^pow * 
    {{{4 * s^2} / {pi * rhoF}} * m}^{2/3}
}

Euler <- function(initial.vals, FPDen, rhoW.at.depth, max.depth = -200000,
                  print.progress = FALSE, convert.units = TRUE){
  #-----------------------------------------------------------------------------
  # Solve equations defining faecal pellet sinking/degradation by integrating with
  # the forward Euler method.
  # Inputs - initial.vals: dataframe defined by User
  #          FPDen: density of fecal pellet
  #          rhoW.at.depth: function defining seawater density continuously over depth
  #          max.depth: depth limit of integration, seafloor or sample depth
  #-----------------------------------------------------------------------------
  output <- initial.vals
  stopLoop <- FALSE # TRUE stops the while loop
  t <- 0 # time [s]
  j <- 0 # iteration counter
  while(!stopLoop){
    t <- t + dt # each loop iteration steps forward one time increment
    j <- j + 1
    distance <- output$speed[j] * dt # distance [cm] traveled in time dt
    depth <- output$depth[j] - distance # new depth
    rD <- rD_fun(-depth) # degradation rate over time step
    mass.loss <- rD * output$mass.total[j] * dt # mass lost (from all faecal pellets) in time dt
    M <- output$mass.total[j] - mass.loss # new mass of all faecal pellets
    new.fragments <- rF * output$number[j] * dt # number of new pieces from fragmentation in time dt
    N <- output$number[j] + new.fragments # new number of faecal pellets
    m <- M / N # mass of individual faecal pellet
    V <- m / FPDen # new volume -- we don't need this to solve the equations, but it may be useful once plastics are considered...
    rho.W <- rhoW.at.depth(depth) # seawater density
    sp <- speed(m, FPDen, rho.W) # updated speed
    output[j+1,] <- data.frame(time = t, depth = depth, number = N, mass.total = M, volume = V, mass = m, speed = sp, rD = rD)
    # Conditions for breaking out of the while loop...
    depth.range.exceeded <- output$depth[j+1] < max.depth # has pellet descended below depth limit of measurements?
    too.slow <- output$speed[j+1] * 86400 < 1 # is pellet moving so slow that it's effectively stationary? (I've chosen 1 cm/day, but this can be changed).
    if(depth.range.exceeded | too.slow) stopLoop <- TRUE
    if(print.progress) print(paste0('iteration ', j))
  }
  #----------------------
  # Integration complete
  #----------------------
  #----------------------
  # Convert units from g-cm-s into something more easily readable
  if(convert.units){
    output$time <- output$time / 86400 # [day]
    output$depth <- output$depth / 100 # [m]
    output$volume <- output$volume * 1000 # [mm^3]
    output$mass <- output$mass * 1000 # [mg]
    output$speed <- output$speed * 864 # [m/day]
    output$rD <- output$rD * 86400 # [1/day]
  }
  return(output)
}
```

Data for average uncontaminated FP (Nov-Feb)
```{r, include=FALSE}
V.init <- volume_fun(L.init, D.init, shape = 'cylinder') # initial volume [cm^3]
m.init <- mass(V.init, gd) # initial mass [g]
S <- shape(L.init, D.init) # shape parameter describing faecal pellets, assumed constant [unitless]

depth.init <- sapply(1:nmonths, function(z) depth.range$min[z]) # initial depths [cm], one for each month, depend on depth range of samples
rhoW.init <- sapply(1:nmonths, function(z) rho.interp[[z]](depth.init[z])) # initial seawater density

sp.init <- speed(m.init, gd, rhoW.init) # initial speed [cm/s]

rD.init <- rD_fun(-depth.init)

N.init <- 1 # initial number of faecal pellets
M.init <- m.init * N.init # initial mass of faecal pellets combined

# Create list of data frames to store output for each month, and populate with initial values
initial.vals <- setNames(lapply(1:nmonths, function(z){
  data.frame(time = 0, depth = depth.init[z], number = N.init, mass.total = M.init,
             volume = V.init, mass = m.init, speed = sp.init[z], rD = rD.init[z])}),
  month.abb[months])

# Call the integrating function to generate output for each month
outputNorm <- lapply(months, function(z){
  mon <- month.abb[z]
  rhoW.at.depth <- rho.interp[[mon]]
  max.depth <- depth.range$max[depth.range$month == z]
  out <- Euler(initial.vals[[mon]], gd, rhoW.at.depth, max.depth)
  out <- cbind(month = mon, out)
  return(out)
})

```

^^^
```{r}
# View results
i <- 1
rbind(head(outputNorm[[i]], 5),
      tail(outputNorm[[i]], 5))

plot(outputNorm[[i]]$time, outputNorm[[i]]$depth, type = 'l',
     xlab = 'time (days)', ylab = 'depth (m)')
plot(outputNorm[[i]]$number, outputNorm[[i]]$depth, type = 'l',
     xlab = expression(number ~ of ~ faecal ~ pellets), ylab = 'depth (m)', log = 'x')
plot(outputNorm[[i]]$mass, outputNorm[[i]]$depth, type = 'l',
     xlab = expression(faecal ~ pellet ~ mass ~ (mg)), ylab = 'depth (m)')
plot(outputNorm[[i]]$mass.total, outputNorm[[i]]$depth, type = 'l',
     xlab = expression(total ~ mass ~ (mg)), ylab = 'depth (m)')
plot(outputNorm[[i]]$speed, outputNorm[[i]]$depth, type = 'l',
     xlab = 'sinking speed (m/day)', ylab = 'depth (m)')
plot(outputNorm[[i]]$volume, outputNorm[[i]]$depth, type = 'l',
     xlab = expression(faecal ~ pellet ~ volume ~ (mm^3)), ylab = 'depth (m)')

# mtext(month.name[months[i]], outer = TRUE, line = -1.5)



```

Function for calculating mass of MP contaminated FP
```{r}
# Function for calculating mass of MP contaminated FP
# Inputs - nMP: # of MP particles
# fibRatio: proportion MP that are fibers (assuming other MPs are all fragments)
# fibV: Volume of fiber
# fragV: Volume of fragment
# FPVol: Initial volume of FP
# FPDens: Density of uncontamintaed FP
# polyrats: User-defined DF of plastic properties and proportions
FPMass.2 <- function(nMP, fibRatio, fibV, fragV, FPVol, FPDens, polyrats){
  fibK <- fibRatio * nMP
  fragK <- (1 - fibRatio) * nMP
  MPVol <-  (fibK * fibV) + (fragK * fragV)
  ratio <- MPVol/FPVol
  Ovrll <- sum(polyrats$plymrs * polyrats$ratios) # average polymer density
  Fd <- (FPDens * (1-ratio)) + (Ovrll * ratio)
  FP.init <- mass(FPVol, Fd)
  return(c(mass = FP.init, density = Fd)) # Best practice is to define function output
}

# Same function as previous, but polyrats is substituted for a known density
FPMass.3 <- function(nMP, fibRatio, fibV, fragV, FPVol, FPDens, PolyDens){
  fibK <- fibRatio * nMP
  fragK <- (1 - fibRatio) * nMP
  MPVol <-  (fibK * fibV) + (fragK * fragV)
  ratio <- MPVol/FPVol
  Fd <- (FPDens * (1-ratio)) + (PolyDens * ratio)
  FP.init <- mass(FPVol, Fd)
  return(c(mass = FP.init, density = Fd)) # Best practice is to define function output
}
```

^^^
```{r, include=FALSE}
# Get MP fiber and fragment volumes
fib <- volume_fun(0.013, 0.003, plastic.type = 'fibre') # WHERE ARE THESE NUMBERS COMING FROM? INCLUDE SOURCE IN THE COMMENTS. ENSURE THE UNITS ARE IN cm.
frag <- volume_fun(0.002, 0.00142, plastic.type = 'fragment')

# create df with polymer density properties and proportions
polyrats <- data.frame(name = c("PA6", "HDPE", "VTE", "PBR"),
                       plymrs = c(1.14,0.95,0.93,0.9),
                       ratios = c(0.55,0.18,0.17,0.1))

FPMass <- FPMass.2(2.13, 0.82, fib, frag, V.init, gd, polyrats)

# sp.init <- speed(FPMass[1], FPMass[2], rhoW.init) # initial speed [cm/s]
sp.init <- speed(FPMass['mass'], FPMass['density'], rhoW.init) # initial speed [cm/s]

M.init <- FPMass['mass'] * N.init

# Create list of data frames to store output for each month, and populate with initial values
initial.vals <- setNames(lapply(1:nmonths, function(z){
  data.frame(time = 0, depth = depth.init[z], number = N.init, mass.total = M.init, volume = V.init, mass = FPMass['mass'], speed = sp.init[z], rD = rD.init[z])}),
  month.abb[months])

outputMP <- lapply(months, function(z){
  mon <- month.abb[z]
  rhoW.at.depth <- rho.interp[[mon]]
  max.depth <- depth.range$max[depth.range$month == z]
  out <- Euler(initial.vals[[mon]], FPMass['density'], rhoW.at.depth, max.depth)
  out <- cbind(month = mon, out)
  return(out)
})

```

^^^
```{r}
i <- 1
rbind(head(outputMP[[i]], 5),
      tail(outputMP[[i]], 5))

plot(outputMP[[i]]$time, outputMP[[i]]$depth, type = 'l',
     xlab = 'time (days)', ylab = 'depth (m)')
plot(outputMP[[i]]$mass, outputMP[[i]]$depth, type = 'l',
     xlab = expression(faecal ~ pellet ~ mass ~ (mg)), ylab = 'depth (m)')
plot(outputMP[[i]]$speed, outputMP[[i]]$depth, type = 'l',
     xlab = 'sinking speed (m/day)', ylab = 'depth (m)')
plot(outputMP[[i]]$volume, outputMP[[i]]$depth, type = 'l',
     xlab = expression(faecal ~ pellet ~ volume ~ (mm^3)), ylab = 'depth (m)')
mtext(month.name[months[i]], outer = TRUE, line = -1.5)
```

MP 2050 scenario
```{r, include=FALSE}
# Get MP fiber and fragment volumes
fib <- volume(0.013, 0.003)
frag <- volume(0.002,0.00142)

# create df with polymer density properties and proportions
polyrats <- data.frame(name = c("PA6", "HDPE", "VTE", "PBR"),
                       plymrs = c(1.14,0.95,0.93,0.9),
                       ratios = c(0.55,0.18,0.17,0.1))

FPMass <- FPMass.2(8.52, 0.82, fib, frag, V.init, gd, polyrats)

sp.init <- speed(FPMass['mass'], FPMass['density'], rhoW.init) # initial speed [cm/s]

# Create list of data frames to store output for each month, and populate with initial values
output <- setNames(lapply(1:nmonths, function(z){
  data.frame(time = 0, depth = depth.init[z], volume = V.init, mass = FPMass['mass'], speed = sp.init[z], rD = rD.init[z])}),
  month.abb[months])

outputMP2 <- Euler(output, FPMass['density'])
```

```{r}
i <- 1
rbind(head(outputMP2[[i]], 5),
      tail(outputMP2[[i]], 5))

plot(outputMP2[[i]]$time, outputMP2[[i]]$depth, type = 'l',
     xlab = 'time (days)', ylab = 'depth (m)')
plot(outputMP2[[i]]$mass, outputMP2[[i]]$depth, type = 'l',
     xlab = expression(faecal ~ pellet ~ mass ~ (mg)), ylab = 'depth (m)')
plot(outputMP2[[i]]$speed, outputMP2[[i]]$depth, type = 'l',
     xlab = 'sinking speed (m/day)', ylab = 'depth (m)')
plot(outputMP2[[i]]$volume, outputMP2[[i]]$depth, type = 'l',
     xlab = expression(faecal ~ pellet ~ volume ~ (mm^3)), ylab = 'depth (m)')
mtext(month.name[months[i]], outer = TRUE, line = -1.5)
```

```{r}
tail(outputNorm[[1]])
tail(outputMP[[1]])
tail(outputMP2[[1]])
```

```{r}
# Set sequences for plastic concentration (# of pieces per individual) and 
# plastic density (overall density of all plastic polymers)
polCon <- seq(0,100,1)
polDen <- seq(0.9,1.15,0.01)

# Set test values in quantile proportions
polCon2 <- as.vector(c(quantile(polCon, 0.25), quantile(polCon, 0.5), quantile(polCon, 0.75), quantile(polCon, 1))) # quartiles without 0
polDen2 <- as.vector(c(quantile(polDen, 0), quantile(polDen, 0.33), quantile(polDen, 0.66), quantile(polDen, 1))) # in thirds

# All combination
ConDen2 <- expand.grid(pCon = polCon2, pDen = polDen2)
```

```{r}
# line defaults to "ragged" despite "shape" being reset in code
# fib2 <- volume_fun(130,plastic.type = 'fibre')
# volume of fragment based on L, W and A
frag2 <- suppressWarnings(volume_fun(20,14.2, 284))
frag2 <- frag2 * 10^-12 # convert from micron^3 to cm^3

outputAll <- list()
par(mfrow=c(2,2))
for(i in 1:nrow(ConDen2)){
  FPMass <- FPMass.3(ConDen2$pCon[i], 0.82, fib, frag2, V.init, gd, ConDen2$pDen[i])
  sp.init <- speed(FPMass['mass'], FPMass['density'], rhoW.init) # initial speed [cm/s]
  
  output <- setNames(lapply(1:nmonths, function(z){
    data.frame(time = 0, depth = depth.init[z], volume = V.init, mass = FPMass['mass'], speed = sp.init[z], rD = rD.init[z], month = months[z])}), month.abb[months])
  
  outputMain <- Euler(output, FPMass['density'])
  new.Main <- do.call('rbind', outputMain)
  new.Main <- new.Main %>% mutate(PolyDen = as.factor(ConDen2$pDen[i]), PolyCon = as.factor(ConDen2$pCon[i]))
  outputAll[[i]] <- new.Main
}
```
Base R plots
```{r}
m <- 11 # designate month to pull data from
par(mfrow=c(2,2))
for(i in 1:length(outputAll)){
  plot(depth ~ time, data = subset(outputAll[[i]], month == m), type = 'l',
       xlab = 'time (days)', ylab = 'depth (m)')
}

for(i in 1:length(outputAll)){
 plot(depth ~ mass, data = subset(outputAll[[i]], month == m), type = 'l',
    xlab = expression(faecal ~ pellet ~ mass ~ (mg)), ylab = 'depth (m)')
}

for(i in 1:length(outputAll)){
  plot(depth ~ speed, data = subset(outputAll[[i]], month == m), type = 'l',
       xlab = 'sinking speed (m/day)', ylab = 'depth (m)')
}

for(i in 1:length(outputAll)){
  plot(depth ~ volume, data = subset(outputAll[[i]], month == m), type = 'l',
       xlab = expression(faecal ~ pellet ~ volume ~ (mm^3)), ylab = 'depth (m)')
}
```

```{r}
# set limits for DF in list
k <- 1
l <- 4 # chose 4 for polDen and polCon vector lengths (thus possible combinations)
APPA <- list()
# for loop combines DFs in groups of 4, to reflect possible combinations
for(i in 1:4){
  new.All <- do.call('rbind', outputAll[k:l])
  APPA[[i]] <- new.All
  k <- k + 4
  l <- l + 4
}
tail(APPA[[3]])
```
```{r}
m <- 11 # designate month to pull data from
MonthDat <- subset(APPA[[1]], month == m)

# Plot ranges togeter by plastic concentration
ggplot(MonthDat, aes(time, depth, color = PolyCon)) + geom_line()
ggplot(MonthDat, aes(mass, depth, color = PolyCon)) + geom_line()
ggplot(MonthDat, aes(volume, depth, color = PolyCon)) + geom_line()
ggplot(MonthDat, aes(speed, depth, color = PolyCon)) + geom_line()
```

